Rates and delay times of type Ia supernovae in the Dark Energy Survey

### 2021-04-21 draft.tex
Updates from Overleaf

SNe Ia do not occur instantaneously after an episode of star formation. Instead stars with zero-age main-sequence (ZAMS) masses less than $8~\mathrm{M}_{\odot}$ take time to evolve along the main sequence and form WDs and this time is not well constrained. In many works this ``prompt time" which we denote $t_{\mathrm{p}}$ is fixed to some value expected to be the minimum time for a star to evolve off the main sequence and become a WD, such as 40~Myr \citep{Maoz2012,Graur2013,Graur2014}. Other studies have left $t_{\mathrm{p}}$ as a free parameter \citep{Heringer2019,Castrillo2020}. We perform three fits: one with $t_{\mathrm{p}}$ fixed at 40 Myr, one with it as a free parameter along with $\beta$ and $A$, and a third with $\beta$ and $A$ fixed and  $t_{\mathrm{p}}$ free. We fit $t_{\mathrm{p}}$ in log space and adopt a Gaussian prior with hyper-parameters $p(\log(t_{\mathrm{p}})) \sim \mathcal{N}(-1.3\,,0.5)$, which corresponds to being centred around 50~Myr. Choosing this regime to fit allows more prior weight to be placed on shorter prompt times as per the majority of the literature. We use the $\hat{R}$ diagnostic of \citep{Vehtari2019} to assess the convergence of MC chains. When including $t_{\mathrm{p}}$ as a free parameter, we find that the strong degeneracy between $t_{\mathrm{p}}$ and $A$ cause the sampler to fail to converge with $\hat{R} = 1.3$ for the chains of those two parameters. In subsequent analyses we adopt as our fiducial model that with fixed $t_{\mathrm{p}} =0.04$~Gyr.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs/fit_fine_BC03_corner.png}
    \caption{Joint posterior distribution for the slope and intercept of the linear fit to the SN rate per galaxy per year as a function of stellar mass (Section \ref{sec:rates}; Fig. \ref{fig:rate_fitted}).}
    \label{fig:corner_slope_int}
\end{figure}
We model the relation with the linear relationship: 
\begin{equation}
    R_{\mathrm{G}} = \frac{\mathrm{d}R}{\mathrm{d}M_*} M_* + c \,,
\label{eq:rate_fit}
\end{equation}
where $dR/dM_*$ signifies the change of the rate of SNe as a function of the stellar mass, $c$ is a constant that sets the normalisation of the rate. We fit the model assuming a normal likelihood, such that the observed rate in each stellar mass bin is itself modelled as a Gaussian distribution described by the mean and standard deviation of the data in that bin. We adopt weakly informative normal priors on the slope and intercept: $p\left(\frac{\mathrm{d}R}{\mathrm{d}M_*}\right) \sim \mathcal{N}\left(0,5\right)$ and  $p\left(c\right) \sim \mathcal{N}\left(-12,5\right)$ respectively. We sample using 4 chains, each with 2000 warm up and 2000 sampling iterations. We report parameter estimates based on the mean and standard deviation of their posterior samples.

The joint posterior distribution for the slope and intercept of the overall SN rate vs stellar mass are shown in Fig. \ref{fig:corner_slope_int}. The two parameters are highly degenerate, yet well constrained. This degeneracy manifests as the spread of potential linear fits as drawn in light blue on Fig. \ref{fig:rate_fitted}.

%\section{Changes to the nominal SFH model}
%\label{appendix:model_adjustments}
%In Section \ref{sec:model} we describe our toy model of galaxy SFHs based on the work of %\citetalias{Childress2014}. Here we

>>>>

SNe Ia do not occur instantaneously after an episode of star formation. Instead stars with zero-age main-sequence (ZAMS) masses less than $8~\mathrm{M}_{\odot}$ take time to evolve along the main sequence and form WDs and this time is not well constrained. In many works this ``prompt time" which we denote $t_{\mathrm{p}}$ is fixed to some value expected to be the minimum time for a star to evolve off the main sequence and become a WD, such as 40~Myr \citep{Maoz2012,Graur2013,Graur2014}. Other studies have left $t_{\mathrm{p}}$ as a free parameter \citep{Heringer2019,Castrillo2020}. We perform three fits: one with $t_{\mathrm{p}}$ fixed at 40 Myr, one with it as a free parameter along with $\beta$ and $A$, and a third with $\beta$ and $A$ fixed and  $t_{\mathrm{p}}$ free. We fit $t_{\mathrm{p}}$ in log space and adopt a Gaussian prior with hyper-parameters $p(\log(t_{\mathrm{p}})) \sim \mathcal{N}(-1.3\,,0.5)$, which corresponds to being centred around 50~Myr. Choosing this regime to fit allows more prior weight to be placed on shorter prompt times as per the majority of the literature. We use the $\hat{R}$ diagnostic of \citep{Vehtari2019} to assess the convergence of MC chains. When including $t_{\mathrm{p}}$ as a free parameter, we find that the strong degeneracy between $t_{\mathrm{p}}$ and $A$ cause the sampler to fail to converge with $\hat{R} \sim 1.3$ for the chains of those two parameters. In subsequent analyses we adopt as our fiducial model that with fixed $t_{\mathrm{p}} =0.04$~Gyr.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs/fit_fine_BC03_corner.png}
    \caption{Joint posterior distribution for the slope and intercept of the linear fit to the SN rate per galaxy per year as a function of stellar mass (Section \ref{sec:rates}; Fig. \ref{fig:rate_fitted}).}
    \label{fig:corner_slope_int}
\end{figure}
We model the relation with the linear relationship: 
\begin{equation}
    R_{\mathrm{G}} = \frac{\mathrm{d}R}{\mathrm{d}M_*} M_* + c \,,
\label{eq:rate_fit}
\end{equation}
where $dR/dM_*$ signifies the change of the rate of SNe as a function of the stellar mass, $c$ is a constant that sets the normalisation of the rate. We fit the model assuming a normal likelihood, such that the observed rate in each stellar mass bin is itself modelled as a Gaussian distribution described by the mean and standard deviation of the data in that bin. We adopt weakly informative normal priors on the slope and intercept: $p\left(\frac{\mathrm{d}R}{\mathrm{d}M_*}\right) \sim \mathcal{N}\left(0,5\right)$ and  $p\left(c\right) \sim \mathcal{N}\left(-12,5\right)$ respectively. We sample using 4 chains, each with 2000 warm up and 2000 sampling iterations. We report parameter estimates based on the mean and standard deviation of their posterior samples. We use the $\hat{R}$ diagnostic of \citep{Vehtari2019} to assess the convergence of MC chains, and only report fits where $\hat{R}\let1.05$

The joint posterior distribution for the slope and intercept of the overall SN rate vs stellar mass are shown in Fig. \ref{fig:corner_slope_int}. The two parameters are highly degenerate, yet well constrained. This degeneracy manifests as the spread of potential linear fits as drawn in light green on Fig. \ref{fig:rate_fitted}.

\section{Posterior distributions for DTD fits}

In this section we present the joint and marginal posterior distributions for the various fits of the DTD presented in Section \ref{sec:split_x1_c}. Posteriors have been displayed using v0.33.0 of \texttt{ChainConsumer} \citep{Hinton2016}.

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/fit_dtd_split_x1_A_beta_corner.png}
    \caption{Posterior distributions for DTD slope $\beta$ and normalisation $A$ for SNe Ia split by their $x_1$ parameter.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:corner_beta_norm_split_x1}}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/fit_dtd_split_c_A_beta_corner.png}
    \caption{Posterior distributions for DTD slope $\beta$ and normalisation $A$ for SNe Ia split by their $c$ parameter.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:corner_beta_norm_split_c}}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/fitjoint_t1_t2_corner.png}
    \caption{Posterior distributions for $t_1$ and $t_2$ of the model presented in Section \ref{subsec:split_x1}.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:corner_t1_t2}}
\end{figure}
